State-space PTA

### 2023-04-06 docs/manuscript/paper.tex
EoD

%%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\kms}{\,km\,s$^{-1}$} % kilometres per second
\newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough

@@

\subsection{Kalman Filtering}\label{sec:kalman_filter}
The Kalman filter \citep{Kalman1} is a algorithmic technique for recovering a set of system state variables $\bar{x}$ given some noisy measurements $\bar{z}$. It is a common technique in signal processing that has also been applied more recently with great success in astrophysics \citep[e.g.][]{Meyers2021,Melatos2023}. The linear Kalman filter operates on measurements that are related to states via a linear transformation
\begin{equation}
	z = Hx + v
\end{equation}
where $H$ is the measurement matrix and $v$ a Gaussian measurement noise. The underlying states are the solutions to the state-space equation 
\begin{equation}
	\dot{x} = Fx + Gu + w
\end{equation}
where $x$ are the state variables, $F$ the system dynamics matrix, $G u$ the control matrix/vector and $w$ a stochastic zero-mean process which .  By comparison with the preceding equations Eqs. 	\ref{eq:state1} - 	\ref{eq:state2} it is immediately obvious how our state space model maps onto the Kalman filter structure. Specifically, our states are just the $N$ intrinsic pulsar frequencies $\bar{x} = (f_1,f_2,...f_N)$ whilst our measurements are the $N$ measured pulse frequencies $\bar{z} = (f^{(M)}_1,f^{(M)}_2,...f^{(M)}_N)$. If we specialize to the case of constant time sampling between our observations, then for our formulation, the various components that make up the Kalman filter are as follows: 
\begin{equation}
	F_i = F_{i+1} = e^{-\gamma \Delta t}
\end{equation}
\begin{align}
	T_i &= \int_{t_i}^{t_{i+1}}  e^{A (t_{i+1} - t')} N(t') dt' \\
	    & =
\end{align}

The Kalman filter includes the effect of process noise and measurement noise via the defines a measurement noise matrix $R = E[v v^T]$
defines a process noise matrix $Q = E[w w^T]$.

\begin{equation}
	\eta_i = \int_{t_i}^{t_{i+1}}  e^{A (t_{i+1} - t')} \sigma dt'
\end{equation}

Whilst for our problem we only need concern ourselves with linear stuff

%Extensions to non-linear state-transitions
%[i.e. 𝑭𝑿 → 𝑓 (𝑿) in equation (11)], and non-linear measurement
%equations [i.e. 𝑪𝑿 → 𝑐(𝑿) in equation (10)] can be achieved using a range of tools such as the extended Kalman filter (Jazwinski
%1970), the unscented Kalman filter (Julier & Uhlmann 1997; Wan
%& Merwe 2000; Julier & Uhlmann 2004), or particle filters (Del
%Moral 1997).
%The full set of Kalma
%
for a full review we refer the reader to X

the state the states are related to a measurement via

where this is that is this

By comaprison with the state space equations we derived earlier (Eqs A to B) we can see that our continuous equations map ideally onto this structure. In particular, for our system our state $\bar{x}$ is simply the frequencies of each individual pulsar $= (f_1, f_2, f_3,...)$

>>>>

%%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%%
\usepackage{subcaption}
% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\kms}{\,km\,s$^{-1}$} % kilometres per second
\newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough

@@

\subsection{Kalman Filtering}\label{sec:kalman_filter}
The Kalman filter \citep{Kalman1} is a algorithmic technique for recovering a set of system state variables $\bar{x}$ given some noisy measurements $\bar{z}$. It is a common technique in signal processing that has also been applied more recently with great success in astrophysics \citep[e.g.][]{Meyers2021,Melatos2023}. The linear Kalman filter operates on measurements that are related to states via a linear transformation
\begin{equation}
	z = Hx + v
\end{equation}
where $H$ is the measurement matrix and $v$ a Gaussian measurement noise. The underlying states are the solutions to the state-space equation 
\begin{equation}
	\dot{x} = Fx + Gu + w
\end{equation}
where $x$ are the state variables, $F$ the system dynamics matrix, $G u$ the control matrix/vector and $w$ a stochastic zero-mean process which .  By comparison with the preceding equations Eqs. 	\ref{eq:state1} - 	\ref{eq:state2} it is immediately obvious how our state space model maps onto the Kalman filter structure. Specifically, our states are just the $N$ intrinsic pulsar frequencies $\bar{x} = (f_1,f_2,...f_N)$ whilst our measurements are the $N$ measured pulse frequencies $\bar{z} = (f^{(M)}_1,f^{(M)}_2,...f^{(M)}_N)$. If we specialize to the case of constant time sampling between our observations, then for our formulation, the components that make up the Kalman filter are as follows: 
\begin{equation}
	F_i = F_{i+1} = e^{-\gamma \Delta t}
\end{equation}
\begin{align}
	T_i &= \int_{t_i}^{t_{i+1}}  e^{A (t_{i+1} - t')} N(t') dt' \\
	    & = f_{\rm EM}(0) + \dot{f}_{\rm EM}(0)  (\Delta t + t_i) - e^{-\gamma \Delta t} (f_{\rm EM}(0) + \dot{f}_{\rm EM}(0)t_i)
\end{align}
\begin{equation}
	H_i = 1 -A(\theta_{\rm GW}) \cos(-\Omega t_i (1 + n\cdot q ) + \Phi_0)
\end{equation}
and $A$ is a constant that is given by  Eq. \ref{eq:state2}. The Kalman filter includes the effect of process noise $w$ and the measurement noise $v$ via the definition of a process noise matrix $Q = E[w w^T]$ and a measurement noise matrix $R = E[v v^T]$, which have the discrete form,
\begin{equation}
	Q_i  \delta_{ij}= \langle \eta_i \eta_j^T \rangle = \frac{- \sigma^2}{2 \gamma} \left( e^{-2 \gamma \Delta t} -1\right)
\end{equation}
\begin{equation}
	R_i = R_{i+1} = \Sigma^2
\end{equation}
For our formulation we are concerned only with the linear Kalman filter since the measurements are a linear function of the states and the state transitions are also linear. Extension to non-linear problems is straighforward using either an extended Kalman filter, the unscented Kalman filter or the particle filter. For a full review of the  Kalman filter equations which may e unfamilar to some, we refer the reader to the appendix X of \cite{Melatos2023}. \newline

\noindent In order for the Kalman filter to function successfully, the parameters of the model, which appear in the various Kalman matrices, must be accurate (e.g. Fig \ref{fig:four figures}). The filter tracks the error in the

The Kalman filte rprovides an estiamte In this way,

\begin{figure*}
	%\centering % Not needed
	\begin{subfigure}[b]{1\columnwidth}
		\includegraphics[width=\textwidth]{images/KF1}
		\caption{}
		\label{fig:6MB_BFS}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{1\columnwidth}
		\includegraphics[width=\textwidth]{images/KF2}
		\caption{}
		\label{fig:25MB_bfs}
	\end{subfigure}
	\label{fig:four figures}
\end{figure*}
